---
title: "Average Frequency of sRNAs"
author: "Carter Newton"
output: html_document
---

### Figure 1 Actual Data
\ 

- create data.frame of original data 
```{r echo=TRUE, results='hide'}
ns <- read.delim("OriginalData/normalized_sRNA.txt")
nm <- read.delim("OriginalData/normalized_mRNA.txt")
```
\ 

- next create data frame pushing column 1 "X" into the row names 
```{r echo=TRUE, results='hide'}
ns_row <- ns
nm_row <- nm
rownames(ns_row) <- ns_row$X
ns_row <- subset(ns_row, select = -X)
rownames(nm_row) <- nm_row$X
nm_row <- subset(nm_row, select = -X)
```
\ 

- remove all genes that contain no sRNA mapping in every sample 
```{r echo=TRUE, results='hide'}
ns_mapping1 <- ns_row[rowSums(ns_row[])>0,]
```
\ 

- create data frames specific to tissue type 
```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(dplyr)
```

```{r echo=TRUE, results='hide'}
ns_mapgeneR <- ns_mapping1 %>% select(ends_with('R'))
ns_mapgeneE <- ns_mapping1 %>% select(ends_with('E'))
ns_mapgeneS <- ns_mapping1 %>% select(ends_with('S'))
ns_mapgeneI <- ns_mapping1 %>% select(ends_with('I'))
ns_mapgeneL <- ns_mapping1 %>% select(ends_with('L'))
```
\ 

- filter out genes with no mapping
```{r echo=TRUE, results='hide'}
ns_mapgeneR <- ns_mapgeneR[rowSums(ns_mapgeneR[])>0,]
ns_mapgeneE <- ns_mapgeneE[rowSums(ns_mapgeneE[])>0,]
ns_mapgeneS <- ns_mapgeneS[rowSums(ns_mapgeneS[])>0,]
ns_mapgeneI <- ns_mapgeneI[rowSums(ns_mapgeneI[])>0,]
ns_mapgeneL <- ns_mapgeneL[rowSums(ns_mapgeneL[])>0,]
```
\ 

- after "summing" out the tissue specific variables, we now will form whole integers by dividing each column by the smallest value in it and then recombining all the columns back together

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
#For endosperm
ns_mapgeneE$row <- 1:nrow(ns_mapgeneE)
ns_mapgeneE$join <- row.names(ns_mapgeneE) #is this needed? may as well keep it
rownames(ns_mapgeneE) <- ns_mapgeneE$row
ns_mapgeneE <- subset(ns_mapgeneE, select = -row)

n <- ncol(ns_mapgeneE)
i <- 2
c <- min(select(ns_mapgeneE, 1)[select(ns_mapgeneE, 1)!=min(select(ns_mapgeneE, 1))])
ns_testE <- select(ns_mapgeneE, 1)/ c
ns_testE <- round(ns_testE)
ns_testE$join <- row.names(ns_testE)
while (i<n) {
  c <- min(select(ns_mapgeneE, i)[select(ns_mapgeneE, i)!=min(select(ns_mapgeneE, i))])
  ns_testEmerge <- select(ns_mapgeneE, i)/ c
  ns_testEmerge <- round(ns_testEmerge)
  ns_testEmerge$join <- row.names(ns_testEmerge)
  ns_testE <- merge(ns_testE, ns_testEmerge, sort = FALSE)
  i <- i+1
}
ns_Esort <- ns_testE


#For Root
ns_mapgeneR$row <- 1:nrow(ns_mapgeneR)
ns_mapgeneR$join <- row.names(ns_mapgeneR) #is this needed? may as well keep it
rownames(ns_mapgeneR) <- ns_mapgeneR$row
ns_mapgeneR <- subset(ns_mapgeneR, select = -row)

n <- ncol(ns_mapgeneR)
i <- 2
c <- min(select(ns_mapgeneR, 1)[select(ns_mapgeneR, 1)!=min(select(ns_mapgeneR, 1))])
ns_testR <- select(ns_mapgeneR, 1)/ c
ns_testR <- round(ns_testR)
ns_testR$join <- row.names(ns_testR)
while (i<n) {
  c <- min(select(ns_mapgeneR, i)[select(ns_mapgeneR, i)!=min(select(ns_mapgeneR, i))])
  ns_testRmerge <- select(ns_mapgeneR, i)/ c
  ns_testRmerge <- round(ns_testRmerge)
  ns_testRmerge$join <- row.names(ns_testRmerge)
  ns_testR <- merge(ns_testR, ns_testRmerge, sort = FALSE)
  i <- i+1
}
ns_Rsort <- ns_testR


#For internode
ns_mapgeneI$row <- 1:nrow(ns_mapgeneI)
ns_mapgeneI$join <- row.names(ns_mapgeneI) #is this needed? may as well keep it
rownames(ns_mapgeneI) <- ns_mapgeneI$row
ns_mapgeneI <- subset(ns_mapgeneI, select = -row)

n <- ncol(ns_mapgeneI)
i <- 2
c <- min(select(ns_mapgeneI, 1)[select(ns_mapgeneI, 1)!=min(select(ns_mapgeneI, 1))])
ns_testI <- select(ns_mapgeneI, 1)/ c
ns_testI <- round(ns_testI)
ns_testI$join <- row.names(ns_testI)
while (i<n) {
  c <- min(select(ns_mapgeneI, i)[select(ns_mapgeneI, i)!=min(select(ns_mapgeneI, i))])
  ns_testImerge <- select(ns_mapgeneI, i)/ c
  ns_testImerge <- round(ns_testImerge)
  ns_testImerge$join <- row.names(ns_testImerge)
  ns_testI <- merge(ns_testI, ns_testImerge, sort = FALSE)
  i <- i+1
}
ns_Isort <- ns_testI


#For shoot
ns_mapgeneS$row <- 1:nrow(ns_mapgeneS)
ns_mapgeneS$join <- row.names(ns_mapgeneS) #is this needed? may as well keep it
rownames(ns_mapgeneS) <- ns_mapgeneS$row
ns_mapgeneS <- subset(ns_mapgeneS, select = -row)

n <- ncol(ns_mapgeneS)
i <- 2
c <- min(select(ns_mapgeneS, 1)[select(ns_mapgeneS, 1)!=min(select(ns_mapgeneS, 1))])
ns_testS <- select(ns_mapgeneS, 1)/ c
ns_testS <- round(ns_testS)
ns_testS$join <- row.names(ns_testS)
while (i<n) {
  c <- min(select(ns_mapgeneS, i)[select(ns_mapgeneS, i)!=min(select(ns_mapgeneS, i))])
  ns_testSmerge <- select(ns_mapgeneS, i)/ c
  ns_testSmerge <- round(ns_testSmerge)
  ns_testSmerge$join <- row.names(ns_testSmerge)
  ns_testS <- merge(ns_testS, ns_testSmerge, sort = FALSE)
  i <- i+1
}
ns_Ssort <- ns_testS


#For leaf
ns_mapgeneL$row <- 1:nrow(ns_mapgeneL)
ns_mapgeneL$join <- row.names(ns_mapgeneL) #is this needed? may as well keep it
rownames(ns_mapgeneL) <- ns_mapgeneL$row
ns_mapgeneL <- subset(ns_mapgeneL, select = -row)

n <- ncol(ns_mapgeneL)
i <- 2
c <- min(select(ns_mapgeneL, 1)[select(ns_mapgeneL, 1)!=min(select(ns_mapgeneL, 1))])
ns_testL <- select(ns_mapgeneL, 1)/ c
ns_testL <- round(ns_testL)
ns_testL$join <- row.names(ns_testL)
while (i<n) {
  c <- min(select(ns_mapgeneL, i)[select(ns_mapgeneL, i)!=min(select(ns_mapgeneL, i))])
  ns_testLmerge <- select(ns_mapgeneL, i)/ c
  ns_testLmerge <- round(ns_testLmerge)
  ns_testLmerge$join <- row.names(ns_testLmerge)
  ns_testL <- merge(ns_testL, ns_testLmerge, sort = FALSE)
  i <- i+1
}
ns_Lsort <- ns_testL
```
\ 

- Now we need to form percentages respective to the original figure
```{r echo=TRUE, results='hide'}
#For endosperm
library(plyr)

totalobvE <- (ncol(ns_Esort)-1)*nrow(ns_Esort)
total0E <- ldply(ns_Esort, function(c) sum(c==0))
total0E <- sum(total0E$V1)
totalE <- totalobvE-total0E

total1E <- ldply(ns_Esort, function(c) sum(c==1))
total1E <- total1E[-c(1),]
total1E <- sum(total1E$V1)

total2to10E <- ldply(ns_Esort, function(c) sum(c<=10))
total2to10E <- total2to10E[-c(1),]
total2to10E <- sum(total2to10E$V1)
total2to10E <- total2to10E-(total0E+total1E)

total11to50E <- ldply(ns_Esort, function(c) sum(c<=50))
total11to50E <- total11to50E[-c(1),]
total11to50E <- sum(total11to50E$V1)
total11to50E <- total11to50E-(total0E+total1E+total2to10E)

total51E <- ldply(ns_Esort, function(c) sum(c>=51))
total51E <- total51E[-c(1),]
total51E <- sum(total51E$V1)

percent1E <- total1E/totalE
percent2to10E <- total2to10E/totalE
percent11to50E <- total11to50E/totalE
percent51E <- total51E/totalE


#For root
totalobvR <- (ncol(ns_Rsort)-1)*nrow(ns_Rsort)
total0R <- ldply(ns_Rsort, function(c) sum(c==0))
total0R <- sum(total0R$V1)
totalR <- totalobvR-total0R

total1R <- ldply(ns_Rsort, function(c) sum(c==1))
total1R <- total1R[-c(1),]
total1R <- sum(total1R$V1)

total2to10R <- ldply(ns_Rsort, function(c) sum(c<=10))
total2to10R <- total2to10R[-c(1),]
total2to10R <- sum(total2to10R$V1)
total2to10R <- total2to10R-(total0R+total1R)

total11to50R <- ldply(ns_Rsort, function(c) sum(c<=50))
total11to50R <- total11to50R[-c(1),]
total11to50R <- sum(total11to50R$V1)
total11to50R <- total11to50R-(total0R+total1R+total2to10R)

total51R <- ldply(ns_Rsort, function(c) sum(c>=51))
total51R <- total51R[-c(1),]
total51R <- sum(total51R$V1)

percent1R <- total1R/totalR
percent2to10R <- total2to10R/totalR
percent11to50R <- total11to50R/totalR
percent51R <- total51R/totalR


#For seedling
totalobvS <- (ncol(ns_Ssort)-1)*nrow(ns_Ssort)
total0S <- ldply(ns_Ssort, function(c) sum(c==0))
total0S <- sum(total0S$V1)
totalS <- totalobvS-total0S

total1S <- ldply(ns_Ssort, function(c) sum(c==1))
total1S <- total1S[-c(1),]
total1S <- sum(total1S$V1)

total2to10S <- ldply(ns_Ssort, function(c) sum(c<=10))
total2to10S <- total2to10S[-c(1),]
total2to10S <- sum(total2to10S$V1)
total2to10S <- total2to10S-(total0S+total1S)

total11to50S <- ldply(ns_Ssort, function(c) sum(c<=50))
total11to50S <- total11to50S[-c(1),]
total11to50S <- sum(total11to50S$V1)
total11to50S <- total11to50S-(total0S+total1S+total2to10S)

total51S <- ldply(ns_Ssort, function(c) sum(c>=51))
total51S <- total51S[-c(1),]
total51S <- sum(total51S$V1)

percent1S <- total1S/totalS
percent2to10S <- total2to10S/totalS
percent11to50S <- total11to50S/totalS
percent51S <- total51S/totalS


#For leaf
totalobvL <- (ncol(ns_Lsort)-1)*nrow(ns_Lsort)
total0L <- ldply(ns_Lsort, function(c) sum(c==0))
total0L <- sum(total0L$V1)
totalL <- totalobvL-total0L

total1L <- ldply(ns_Lsort, function(c) sum(c==1))
total1L <- total1L[-c(1),]
total1L <- sum(total1L$V1)

total2to10L <- ldply(ns_Lsort, function(c) sum(c<=10))
total2to10L <- total2to10L[-c(1),]
total2to10L <- sum(total2to10L$V1)
total2to10L <- total2to10L-(total0L+total1L)

total11to50L <- ldply(ns_Lsort, function(c) sum(c<=50))
total11to50L <- total11to50L[-c(1),]
total11to50L <- sum(total11to50L$V1)
total11to50L <- total11to50L-(total0L+total1L+total2to10L)

total51L <- ldply(ns_Lsort, function(c) sum(c>=51))
total51L <- total51L[-c(1),]
total51L <- sum(total51L$V1)

percent1L <- total1L/totalL
percent2to10L <- total2to10L/totalL
percent11to50L <- total11to50L/totalL
percent51L <- total51L/totalL


#For internode
totalobvI <- (ncol(ns_Isort)-1)*nrow(ns_Isort)
total0I <- ldply(ns_Isort, function(c) sum(c==0))
total0I <- sum(total0I$V1)
totalI <- totalobvI-total0I

total1I <- ldply(ns_Isort, function(c) sum(c==1))
total1I <- total1I[-c(1),]
total1I <- sum(total1I$V1)

total2to10I <- ldply(ns_Isort, function(c) sum(c<=10))
total2to10I <- total2to10I[-c(1),]
total2to10I <- sum(total2to10I$V1)
total2to10I <- total2to10I-(total0I+total1I)

total11to50I <- ldply(ns_Isort, function(c) sum(c<=50))
total11to50I <- total11to50I[-c(1),]
total11to50I <- sum(total11to50I$V1)
total11to50I <- total11to50I-(total0I+total1I+total2to10I)

total51I <- ldply(ns_Isort, function(c) sum(c>=51))
total51I <- total51I[-c(1),]
total51I <- sum(total51I$V1)

percent1I <- total1I/totalI
percent2to10I <- total2to10I/totalI
percent11to50I <- total11to50I/totalI
percent51I <- total51I/totalI
```
\ 

- with the percentages formed, now i need to get them organized into a data frame for the graph to interpret

```{r echo=TRUE}
library(ggplot2)

figure1A2 <- data.frame('tissue type'=rep(c('Root', 'Endosperm', 'Seedling', 'Internode', 'Leaf'), each=4),
'mapping locations'=rep(c('1', '2to10', '11to50', '>50'), times=5),
'percent'=c(percent1R, percent2to10R, percent11to50R, percent51R, 
            percent1E, percent2to10E, percent11to50E, percent51E,
            percent1S, percent2to10S, percent11to50S, percent51S,
            percent1I, percent2to10I, percent11to50I, percent51I,
            percent1L, percent2to10L, percent11to50L, percent51L))

figure1Agraph <- ggplot(figure1A2, aes(fill=mapping.locations, x=tissue.type, y=percent)) + geom_bar(position = 'stack', stat = 'identity') + coord_flip() + theme_minimal() + scale_fill_brewer(palette = 'YlOrRd')
```
```{r echo=TRUE}
figure1Agraph
```
\ 

- that is all!